---
title: "Freezing Behavior Detection"
output: html_notebook
---
This R-Notebook provides a way to use outputs from DeepLabCut to extract freezing behaviors of the mouse


load needed libraries
```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

import dataset
**change file name as needed
**make sure that the csv file is in the same folder ans this r markdown
```{r message=FALSE, warning=FALSE}
data <- read_csv("test.csv")
```

extract data for the movement of snout
select coordinates for snout only
select only the ones with >=60% likelihood
**change the '1:4' to c(1, srange) if snout is not the first body part, srange is the column indexes containing x, y, and likelihood for snout
```{r}
snout <- data[-c(1:2),1:4]
colnames(snout) <- c("frame", "x", "y", "likelihood")
snout <- snout %>%
  mutate(x = as.numeric(x), y=as.numeric(y)) %>%
  filter(likelihood >= 0.6)
```

add new column of movement of snout between 2 frames and then avg across three frames
```{r}
movement = c()
i <- 1
while(i <= nrow(snout) - 2)
{
  movement[i] <- sqrt(abs(snout[i+1, 2] - snout[i, 2])^2 + abs(snout[i+1, 3] - snout[i, 3]))
  i <- i +1
}
movement[nrow(snout)-1] <- movement[i-1]
movement[nrow(snout)] <- movement[i-1]
movement <- as.numeric(movement)
snout$movement <- movement
i <- 2
while(i <= nrow(snout) - 1)
{
  snout$movement[i] <- (movement[i-1] + movement[i] + movement[i+1]) / 3
  i <- i + 1
}

```


determine freezing frames
change unit to cm/s
mark y/n - movement < 4cm/s (need to calculate according to fps and pixel-per-cm)
**set frame per second to fps
**set pixel per cm to pcm
```{r}
fps <- 20
pcm <- 23
snout$movement <- snout$movement * fps / pcm
snout$move4 <- snout$movement >= 4
```

determine freezing behavior
for loop - slow movement >=1.28s (need fps)
```{r}
period <- as.integer(1.28*fps)
freeze <- c()
i <- 0
r <- 1
v <- r + period - i
while (r <= nrow(snout) - period + 1){
  if (snout$move4[r] == FALSE){
    i <- i + 1
    v <- r + period - i
    if (i < period){
      if(sum(snout$move4[r:v]) == 0){
        freeze[r] <- TRUE
      } else{
        freeze[r] <- FALSE
      }
    } else{
      freeze[r] <- TRUE
    }
  } else{
    i <- 0
    freeze[r] <- FALSE
  }
  r <- r + 1
}

v <- nrow(snout)
while(r <= v){
  if(snout$move4[r] == FALSE){
    i <- i + 1
    if(i < period){
      if(sum(snout$move4[r:v] < 1)){
        freeze[r] <- TRUE
      } else{
        freeze[r] <- FALSE
      }
    } else{
      freeze[r] <- TRUE
    }
  } else{
    freeze[r] <- FALSE
    break
  }
  r <- r + 1
}

if(r < nrow(snout)){
  freeze[r:v] <- FALSE
}

snout$freeze <- freeze
```

store start frame, stop frame, and duration (in seconds) in a separate dataframe
```{r}
start <- c()
stop <- c()
i <- 0
j <- 0
r <- 1
while (r <= nrow(snout)){
  if (snout$freeze[r] == TRUE){
    i <- i + 1
    j <- 0
    if (i == 1){
      start <- append(start, snout$frame[r])
    }
  } else{
    j <- j + 1
    i <- 0
    if (j == 1){
      stop <- append(stop, snout$frame[r-1])
    }
  }
  r <- r + 1
}

if(snout$freeze[r-1]==TRUE){
  stop <- append(stop, snout$frame[r-1])
}

freeze_data <- data.frame('StartFrame' = start, 'StopFrame' = stop) %>%
  mutate(StartFrame=as.numeric(as.character(StartFrame)), StopFrame=as.numeric(as.character(StopFrame))) %>%
  mutate(Duration = StopFrame/fps - StartFrame/fps)

```

output results in csv
**change paths as needed, remember to name the output file
```{r}
write_csv(freeze_data, '/Users/asus/Desktop/URSI/freeze.csv')
```

